---
title: "WP1_data-exploration"
output: html_notebook
date: "2023-04-04"
---

# 1. SETUP

```{r setup}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
# 1. SETUP
# PACKAGES
library(tidyverse)  # process data frame
library(readr)      # process rds, csv 
library(lubridate)  # process time
library(readxl)     # read xlsx

library(lme4)       # fit model
library(AICcmodavg) # calculate AIC
library(effects)    # display effect
#library(lattice)    # visualize data
library(MuMIn)      # compare models
library(INLA)       # INLA model
library(sf)

library(RColorBrewer) # visualize data

# FUNCTION
###centring function
c. <- function(x) {x - mean(x)} #function to standardize/scale 

# rsquared function
```

# 2. LOAD DATA

## Otolith data

```{r}
# 2. LOAD DATA ------------------------------------------------------------

dir_otl <- "D:/OneDrive - UGent/data/WP1/otolith"
otl <- read_rds(file.path(dir_otl, "otl_full.rds"))

# dir to save model file
dir_note <- "./note"

# gis to caclulate area
dir_gis <- "./data/admin"
ices_4bc <- as.data.frame(st_read(file.path(dir_gis, "ices_areas_sub_group_4326_new.gpkg")))
ices_4bc <- ices_4bc %>% 
  filter(Area_27 %in% c("4bc", "7a", "8ab")) %>% 
  rename(IcesAreaGroup.area.km2 = Area_km2,
         IcesAreaGroup = Area_27) %>%
  select(IcesAreaGroup, IcesAreaGroup.area.km2) 

```

```{r}
#data for exploration
# 0 increment, add a small value 0.05 to avoid log issue (smallest increment from ILVO is 0.06)
incremet_0 <- otl %>% filter(AnnulusDiameterIncrement.um == 0) #
otl <- otl %>% mutate(AnnulusDiameterIncrement.um = if_else(AnnulusDiameterIncrement.um == 0, 0.05, AnnulusDiameterIncrement.um))
incremet_small <- otl %>% filter(AnnulusDiameterIncrement.um < 2)

data_otl <- otl %>% 
  left_join(ices_4bc, by = "IcesAreaGroup") 

data_otl <- data_otl %>% mutate(fishid = FishID,
                       increment = AnnulusDiameterIncrement.um,
                       log.increment = log(AnnulusDiameterIncrement.um),
                       age = Age,
                       log.age = log(Age),
                       aac = AgeAtCapture,
                       log.aac = log(AgeAtCapture), 
                       method = OtolithProcessingMethod,
                       datasource = DataSource,
                       pop = IcesAreaGroup,
                       pop.area.km2 = IcesAreaGroup.area.km2,
                       year = GrowingYear,
                       cohort = Cohort,
                       fyear = factor(GrowingYear),
                       fcohort = factor(Cohort),
                       pop.year = paste0(pop, ":", year),
                       pop.cohort = paste0(pop, ":", cohort)
                       # predictors
                       #f = FishingMortality,
                       #ssb.i = SpawningStockBiomass.1000t*1000/IcesAreaGroup.area.km2, #unit: ton/km2
                       )

```

## Temperature data

```{r}
## Temperature data ----
## load and process data
dir_temp <- "./data/temp"

# isimip
isimip <- read_rds(file.path(dir_temp, "isimip_sbt_datras_hist_ssp585.rds")) %>%
  mutate(date = Date,
         year = Year, 
         pop = if_else(IcesArea == "4abc", "4bc", IcesArea),
         temp = as.numeric(isimip_sbt),
         source = "isimip") %>%
  select(date:source) 

# oras5
oras <- read_rds(file.path(dir_temp, "oras5_datras.rds")) %>%
  mutate(date = Date,
         year = Year, 
         pop = if_else(IcesArea == "4abc", "4bc", IcesArea),
         temp = oras_sbt,
         source = "oras5") %>%
  select(date:source)

# nemo-medusa
nm <- read_rds(file.path(dir_temp, "nemomedusa_datras.rds")) %>%
  mutate(date = Date,
         year = as.numeric(Year), 
         pop = if_else(IcesArea == "4abc", "4bc", IcesArea),
         temp = nemomedusa_sbt,
         source = "nemo-medusa") %>%
  select(date:source)

## merge all temp data
# merge and summarize mean(temp) 
data_temp <- bind_rows(isimip, oras, nm) %>%
  filter(pop %in% c("4bc", "7a", "8ab"),
         year >= min(data_otl$year), year <= max(data_otl$year)) %>%
  group_by(source, pop, year) %>%
  summarize(temp = mean(temp))

# get centered temp (c.temp) for each pop
data_temp <- data_temp %>%
  group_by(source, pop) %>%
  mutate(c.temp = c.(temp))
```

## Fishing data

```{r}
## Fishing data ----
# Fishing mortality, Spawning Stock Biomass, Recruitment from ICES Stock Assessment

dir_ices <- "./data/ices"

# sole distribution area - survey datras
datras <- read_sf(file.path(dir_ices, "hl_loc_4abc7a8ab.gpkg"))
datras <- as_tibble(datras) %>%
  mutate(pop = if_else(Area_27 == "4abc", "4bc", Area_27)) %>%
  select(pop, area_km2)

# sole stock assessment
data_sol <- read_rds(file.path(dir_ices, "stock-assessment_2023.rds")) %>%
  rename(ssb = SSB,
         f = `F`) %>%
  select(pop, year, ssb, f, recruitment) %>%
  left_join(datras) %>%
  mutate(ssb.i = ssb/area_km2,
         recruitment.i = recruitment/area_km2)

# plaice stock assessment
data_ple <- read_rds(file.path(dir_ices, "ple_stock-assessment_2023.rds")) %>%
  rename(ssb_ple = SSB,
         recruitment_ple = recruitment) %>%
  select(pop, year, ssb_ple, recruitment_ple) %>%
  left_join(datras) %>%
  mutate(ssb_ple.i = ssb_ple/area_km2,
         recruitment_ple.i = recruitment_ple/area_km2)

```

##Nutrient data

```{r}
## Nutrient data ----
dir_nu <- "./data/nutrient"
data_nu <- read_rds(file.path(dir_nu, "ospar_subset_1978-2017_ices_4abc.rds"))
# summarize all river by year 
data_nu <- data_nu %>% 
  group_by(IcesArea, year) %>% 
  summarize(TN = sum(TN)/1000,
            TP = sum(TP)/1000) %>%
  mutate(IcesArea = "4bc") %>%
  rename(pop = IcesArea)
# unit: 1000tN/year, 1000tP/year
```

# 3. EXPLORE DATA

```{r}
# LOAD SUPPORT FILES AND PACKAGES 
source("./ref/HighstatLib.r") # support
```

```{r}
data <- data_otl
```

Exploration protocol

\# Exploration following Zuur et al., 2010 protocol - <https://besjournals.onlinelibrary.wiley.com/doi/10.1111/j.2041-210X.2009.00001.x>

Protocol

1\. Outlier Y and X (Y is response variable and X is predictor) - boxplot, cleveland dotplot

2\. Homogeneity Y - conditional boxplot

3\. Normality

4\. Zero trouble (ratio has this problem?)

5\. Collinearity X - VIR or scatter plots

6\. Relationship Y and X - (multipanel) scatterplots

7\. Interaction - coplots

8\. Independence Y - variogram

## 1. Outlier

```{r}
otl_ilvo <- otl %>% filter(DataSource == "ILVO")
otl_wur <- otl %>% filter(DataSource == "WUR")

ggplot() +
  geom_point(data = otl_ilvo, aes(x = Length.mm, y = AnnulusDiameter.um), alpha = 0.5, color = "grey") + 
  geom_point(data = otl_wur, aes(x = Length.mm, y = AnnulusDiameter.um), alpha = 0.5, color = "black") + 
  facet_grid(~ Age)
```

### ILVO data

```{r}
# ilvo
## otoliths with big growth
### age5 > 5000
otl_diameter_age5_gte5000 <- otl_ilvo %>% filter(Age == 5, AnnulusDiameter.um >= 5000)
otl_diameter_age5_gte5000 <- otl_ilvo %>% filter(FishID %in% otl_diameter_age5_gte5000$FishID)

### age3 > 4700
otl_diameter_age3_gte4700 <- otl_ilvo %>% filter(Age == 3, AnnulusDiameter.um >= 4700)
otl_diameter_age3_gte4700 <- otl_ilvo %>% filter(FishID %in% otl_diameter_age3_gte4700$FishID)

## otoliths with small growth
#### age4 < 2100
otl_diameter_age4_lt2100 <- otl_ilvo %>% filter(Age == 4, AnnulusDiameter.um < 2100)
otl_diameter_age4_lt2100 <- otl_ilvo %>% filter(FishID %in% otl_diameter_age4_lt2100$FishID)

### age9 < 3300
otl_diameter_age9_lt3300 <- otl_ilvo %>% filter(Age == 9, AnnulusDiameter.um < 3300)
otl_diameter_age9_lt3300 <- otl_ilvo %>% filter(FishID %in% otl_diameter_age9_lt3300$FishID)

# list of otl_ilvo to be checked
otl_ilvo_check <- bind_rows(otl_diameter_age5_gte5000, 
                            otl_diameter_age3_gte4700, 
                            otl_diameter_age4_lt2100, 
                            otl_diameter_age9_lt3300) 

ggplot() +
  geom_point(data = otl_ilvo, aes(x = Length.mm, y = AnnulusDiameter.um), alpha = 0.5, color = "grey") + 
  geom_point(data = otl_wur, aes(x = Length.mm, y = AnnulusDiameter.um), alpha = 0.5, color = "black") + 
  geom_point(data = otl_ilvo_check, aes(x = Length.mm, y = AnnulusDiameter.um), color = "red", alpha = 0.5) + 
  facet_grid(~ Age)
```

All suspected outliers were checked and confirmed not measuring errors.

### WUR data

```{r}
# wur
## otoliths with big growth
### 1 otolith with AnnulusDiameter.um > 5200 at age 7
otl_diameter_age7_gte5200 <- otl_wur %>% filter(Age == 7, AnnulusDiameter.um >= 5200)
otl_diameter_age7_gte5200 <- otl_wur %>% filter(FishID %in% otl_diameter_age7_gte5200$FishID)

## otoliths with small growth
### 1 otolith with very small dimater at age 1
otl_diameter_age1_lt200 <- otl_wur %>% filter(AnnulusDiameter.um < 200)
otl_diameter_age1_lt200 <- otl_wur %>% filter(FishID %in% otl_diameter_age1_lt200$FishID)

### 1 otolith with AnnulusDiameter.um < 1100 at age 2
otl_diameter_age2_lt1100 <- otl_wur %>% filter(Age == 2, AnnulusDiameter.um < 1100)
otl_diameter_age2_lt1100 <- otl_wur %>% filter(FishID %in% otl_diameter_age2_lt1100$FishID)

### 2 otoliths with AnnulusDiameter.um < 600 at age 1  
otl_diameter_age1_lt600 <- otl_wur %>% filter(Age == 1, AnnulusDiameter.um < 600)
otl_diameter_age1_lt600 <- otl_wur %>% filter(FishID %in% otl_diameter_age1_lt600$FishID)


# list of otl_wur to be checked
otl_wur_check <- bind_rows(otl_diameter_age1_lt200, 
                           otl_diameter_age2_lt1100, 
                           otl_diameter_age1_lt600, 
                           otl_diameter_age7_gte5200) 

ggplot() +
  geom_point(data = otl_ilvo, aes(x = Length.mm, y = AnnulusDiameter.um), alpha = 0.5, color = "grey") +
  geom_point(data = otl_wur, aes(x = Length.mm, y = AnnulusDiameter.um), alpha = 0.5, color = "black") + 
  #geom_point(data = otl_ilvo_check, aes(x = Length.mm, y = AnnulusDiameter.um), color = "red", alpha = 0.5) + 
  geom_point(data = otl_wur_check, aes(x = Length.mm, y = AnnulusDiameter.um), color = "blue", alpha = 0.5) +
  facet_grid(~ Age)

```

WUR: except for 1 otolith with very small increment at age 1 (\< 200nm), other suspected wur otoliths having growth curves within the range of all sample. Therefore, only remove the otolith with very small increment at age 1 (FishID: sol_fab_0575)

### Increment and predictors

```{r}
# create data id
data$id <- seq(1,nrow(data),1)

# INCREMENT
ggplot(data = data, aes(x = increment, y = id)) + geom_point()
```

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

# INCREMENT BY AGE
ggplot(data = data %>% filter(age == 1), aes(x = increment, y = id)) + geom_point()
ggplot(data = data %>% filter(age == 2), aes(x = increment, y = id)) + geom_point()
ggplot(data = data %>% filter(age == 3), aes(x = increment, y = id)) + geom_point()
ggplot(data = data %>% filter(age == 4), aes(x = increment, y = id)) + geom_point()
ggplot(data = data %>% filter(age == 5), aes(x = increment, y = id)) + geom_point()
ggplot(data = data %>% filter(age == 6), aes(x = increment, y = id)) + geom_point()
#View(data %>% filter(FishID == filter(data, age == 6, increment > 600)$FishID))
# 1 increment > 600 in age 6 (smartlab number 198980 (7a 2009)) but not a measurement error
ggplot(data = data %>% filter(age == 7), aes(x = increment, y = id)) + geom_point()
ggplot(data = data %>% filter(age == 8), aes(x = increment, y = id)) + geom_point()
ggplot(data = data %>% filter(age == 9), aes(x = increment, y = id)) + geom_point()
#View(data %>% filter(FishID == filter(data, age == 9, increment > 300)$FishID))
# 1 increment > 300 in age 9 (smartlab number 183726 (7a 2007)) but not a measurement error
ggplot(data = data %>% filter(age == 10), aes(x = increment, y = id)) + geom_point()
ggplot(data = data %>% filter(age == 11), aes(x = increment, y = id)) + geom_point()
ggplot(data = data %>% filter(age == 12), aes(x = increment, y = id)) + geom_point()
ggplot(data = data %>% filter(age == 13), aes(x = increment, y = id)) + geom_point()
ggplot(data = data %>% filter(age == 14), aes(x = increment, y = id)) + geom_point()
#View(data %>% filter(FishID == filter(data, age == 14, increment > 150)$FishID))
#View(data %>% filter(FishID == filter(data, age == 14, increment < 1)$FishID))
# 1 increment > 150 in age 14 (smartlab number 201911210486 (7a 2019)) but not a measurement error
ggplot(data = data %>% filter(age == 15), aes(x = increment, y = id)) + geom_point()
ggplot(data = data %>% filter(age == 16), aes(x = increment, y = id)) + geom_point()
ggplot(data = data %>% filter(age == 17), aes(x = increment, y = id)) + geom_point()
#View(data %>% filter(FishID == filter(data, age == 17, increment > 100)$FishID))
# 1 increment > 100 in age 17 (smartlab number 2018042910309 (7a 2018)) but not a measurement error
ggplot(data = data %>% filter(age == 18), aes(x = increment, y = id)) + geom_point()
ggplot(data = data %>% filter(age == 19), aes(x = increment, y = id)) + geom_point()
ggplot(data = data %>% filter(age == 20), aes(x = increment, y = id)) + geom_point()
ggplot(data = data %>% filter(age == 21), aes(x = increment, y = id)) + geom_point()
ggplot(data = data %>% filter(age == 22), aes(x = increment, y = id)) + geom_point()

```

There is unlikely outlier in increment and predictors.

## 2. Homogeneity Y

Check for homogeneity in population and year

There is homogeneity in year after account for year random effect (intercept/age slope)

```{r}
# YEAR
m <- lmer(log(increment) ~ log.age + log.aac + (1 | fishid) + (1 | pop), data, REML = T)
m.year <- lmer(log(increment) ~ log.age + log.aac + (1 | fishid) + (1 | pop) + (1 | pop.year), data, REML = T)
m.year.age <- lmer(log(increment) ~ log.age + log.aac + (1 | fishid) + (1 | pop) + (1 + log.age | pop.year), data, REML = T)

# residual plot
data$res <- resid(m)
data$res.year <- resid(m.year)
data$res.year.age <- resid(m.year.age)

ggplot(data = data, aes(x = fyear, y = res)) + geom_boxplot()
ggplot(data = data, aes(x = fyear, y = res.year)) + geom_boxplot()
ggplot(data = data, aes(x = fyear, y = res.year.age)) + geom_boxplot()

# adding random year effect (intercept/intercept + age slope) improve residual homogeneity 

```

## 3. Normality.

Data is normal distribution. The distribution is skewed especially in age \> 9

Residual plot confirms normal distribution

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = data, aes(x = increment)) + geom_histogram()
ggplot(data = data %>% filter(age %in% c(1,2,3)), aes(x = increment)) + geom_histogram() + facet_grid(~ age)
ggplot(data = data %>% filter(age %in% c(4,5,6)), aes(x = increment)) + geom_histogram() + facet_grid(~ age)
ggplot(data = data %>% filter(age %in% c(7,8,9)), aes(x = increment)) + geom_histogram() + facet_grid(~ age)
ggplot(data = data %>% filter(age %in% c(10,11,12)), aes(x = increment)) + geom_histogram() + facet_grid(~ age)
ggplot(data = data %>% filter(age %in% c(13,14,15)), aes(x = increment)) + geom_histogram() + facet_grid(~ age)

# distribution is normal, the distribution is skewed especially in higher ages > 9

# residual plot
m <- lmer(log(increment) ~ c.(log(age)) + c.(log(aac)) + (1|FishID) + (1 + c.(log(age))|pop.year), data, REML = T)

data$res <- resid(m)
ggplot(data = data, aes(x = res)) + geom_histogram(bins = 100)

# residual plot confirm normal distribution

```

## 4. Zero trouble

Increment data has no 0 problem

## 5. Collinearity X - VIR or scatter plots

There is no collinearity (corvif \< 3)

### isimip

```{r}
data <- data_otl %>% 
  left_join(filter(data_temp, source == "isimip"),
            by = join_by(pop, year)) %>%
  filter(is.na(c.temp) == F) %>%
  left_join(data_sol, by = join_by(pop, year)) %>%
  left_join(data_nu, by = join_by(pop, year))
```

```{r echo=FALSE}
MyVar <- c("log.age", "log.aac", "c.temp", "ssb.i", "recruitment.i", "TN")
corvif(data[,MyVar]) 

MyVar <- c("log.age", "log.aac", "c.temp", "ssb.i", "recruitment.i", "TP")
corvif(data[,MyVar]) 
# no collinearity
```

### oras5

```{r}
data <- data_otl %>% 
  left_join(filter(data_temp, source == "oras5"),
            by = join_by(pop, year)) %>%
  filter(is.na(c.temp) == F) %>%
  left_join(data_sol, by = join_by(pop, year)) %>%
  left_join(data_nu, by = join_by(pop, year))
```

```{r echo=FALSE}
MyVar <- c("log.age", "log.aac", "c.temp", "ssb.i", "recruitment.i", "TN")
corvif(data[,MyVar]) 

MyVar <- c("log.age", "log.aac", "c.temp", "ssb.i", "recruitment.i", "TP")
corvif(data[,MyVar]) 
# no collinearity
```

### nemo-medusa

```{r}
data <- data_otl %>% 
  left_join(filter(data_temp, source == "nemo-medusa"),
            by = join_by(pop, year)) %>%
  filter(is.na(c.temp) == F) %>%
  left_join(data_sol, by = join_by(pop, year)) %>%
  left_join(data_nu, by = join_by(pop, year))
```

```{r echo=FALSE}
MyVar <- c("log.age", "log.aac", "c.temp", "ssb.i", "recruitment.i", "TN")
corvif(data[,MyVar]) 

MyVar <- c("log.age", "log.aac", "c.temp", "ssb.i", "recruitment.i", "TP")
corvif(data[,MyVar]) 
# no collinearity
```

## 6. Relationship Y and X - (multipanel) scatterplots

expect linear relationship

## 7. Interaction

expect interaction of log.age\*c.temp, c.temp\*ssb.i/recruitment.i/f

## 8. Independence Y

Potential dependence: temporal over years

ACF (autocorrelation function) of raw data and model without year (log(increment) \~ log(age)) shows autocorrelation.

ACF with year iid and ar1 reduces the lag1 correlation (0.4 to 0.2) but did not resolve the autocorrelation.

No clear difference between iid and ar1 model.

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# raw acf - strong autocorrelation
# acf(log(data$increment), main = "ACF")

# residual plot
m <- lmer(log(increment) ~ log.age + log.aac + (1 | fishid) + (1 | pop), data, REML = T)
m.year <- lmer(log(increment) ~ log.age + log.aac + (1 | fishid) + (1 | pop) + (1 | pop.year), data, REML = T)
m.year.age <- lmer(log(increment) ~ log.age + log.aac + (1 | fishid) + (1 | pop) + (1 + log.age | pop.year), data, REML = T)

data$res <- resid(m)
data$res.year <- resid(m.year)
data$res.year.age <- resid(m.year.age)

# plot acf
acf(log(data$increment), main = "raw data")
acf(data$res, main = "model no year")
acf(data$res.year, main = "iid (1|year)")
```

## 
