---
title: "3_report"
output:
  html_document:
    df_print: paged
date: "2023-07-18"
---

BEFORE

-   2_analyze data.R

AFTER

-   WP1/report

# 1. SETUP
```{r message=FALSE, warning=FALSE}
# PACKAGES
library(tidyverse)  # process data frame
library(lme4)       # fit model
library(AICcmodavg) # calculate AIC
library(effects)    # display effect
library(MuMIn)      # compare models
library(sf)         # process geospatial data
library(reactable)  # interactive table
library(RColorBrewer)# color for visualization
library(broom.mixed)# summarize model results
library(writexl)    # write excel
library(patchwork)  # arrange plot
library(sjPlot)     # summarize table
library(knitr)      # print sjPlot table
#library(arm)       # to calculate se - do not load to not override select function

# DEFAULT THEME
theme_set(theme_classic())

# FUNCTION
#centring function
c. <- function (x) {(x - mean(x))} 

# rsquared function
rsquared.glmm=function(modlist) {
  do.call(rbind,lapply(modlist,function(i) {
    if(inherits(i,"merMod") | class(i)=="merLmerTest") {
      VarF=var(as.vector(fixef(i) %*% t(i@pp$X))) 
      VarRand=colSums(do.call(rbind,lapply(VarCorr(i),function(j) j[1])))
      VarResid=attr(VarCorr(i),"sc")^2
      Rm=VarF/(VarF+VarRand+VarResid)
      Rc=(VarF+VarRand)/(VarF+VarRand+VarResid)
      Rsquared.mat=data.frame(Class=class(i),Marginal=Rm,Conditional=Rc,
                              AIC=AIC(update(i,REML=F))) } 
    else { print("Function requires models of class lm, lme, mer, or    merMod") 
    } } ) ) }

# WORKING DIRECTORY 
dir_output <- "./output"
dir_report <- "./report"

# POPULATION NAME
# create df_pop
df_pop <- tibble(IcesAreaGroup = c("4bc", "7a", "8ab"),
                 pop = c("4bc", "7a", "8ab"),
                 pop.name = c("North Sea", "Irish Sea", "Bay of Biscay"))

# color palette 
pal <- (brewer.pal(8, "Spectral")) 
```

# 2. LOAD DATA

## 2.1. Otolith data

```{r message=FALSE, warning=FALSE}
dir_otl <- "./data"
otl <- read_rds(file.path(dir_otl, "otl_full.rds"))

## check 0 increment
incremet_0 <- otl %>% filter(AnnulusDiameterIncrement.um == 0) 

## check number of increment/year
data_sum <- otl %>% 
  group_by(GrowingYear, IcesAreaGroup) %>% 
  summarize(n_increment = n())
# >= 10 increment/year for each population
data_sum_10 <- otl %>% 
  group_by(GrowingYear, IcesAreaGroup) %>% 
  summarize(n_increment = n()) %>%
  filter(n_increment >= 10) %>%
  mutate(pop.year = paste0(IcesAreaGroup, ":", GrowingYear))

## create ssb index by dividing ssb by ices area
dir_gis <- "./data/admin"
ices_4bc <- as.data.frame(st_read(file.path(dir_gis, "ices_areas_sub_group_4326_new.gpkg"),
                                   quiet = TRUE))
ices_4bc <- ices_4bc %>% 
  filter(Area_27 %in% c("4bc", "7a", "8ab")) %>% 
  rename(IcesAreaGroup.area.km2 = Area_km2,
         IcesAreaGroup = Area_27) %>%
  select(IcesAreaGroup, IcesAreaGroup.area.km2) 

## process otl data 
# 1. remove 1 fish 8ab cohort < 1985
# 2. remove 1 wur fish with very small age 1 increment (< 200nm)
# 3. 0 increment, add a small value 0.05 to avoid log issue (smallest increment from ILVO is 0.06)
# 4. add area of ices area

data_otl <- otl %>% 
  filter(FishID != "SOL_52_G1_Z.121_26-03-1990_1602") %>%
  filter(FishID != "sol_fab_0575") %>%
  mutate(AnnulusDiameterIncrement.um = if_else(AnnulusDiameterIncrement.um == 0, 0.05, AnnulusDiameterIncrement.um)) %>%
  left_join(ices_4bc, by = "IcesAreaGroup") 

# 5. rename variables to formulate models easier
data_otl <- data_otl %>% mutate(fishid = FishID,
                       increment = AnnulusDiameterIncrement.um,
                       log.increment = log(AnnulusDiameterIncrement.um),
                       age = Age,
                       log.age = log(Age),
                       aac = AgeAtCapture,
                       log.aac = log(AgeAtCapture), 
                       method = OtolithProcessingMethod,
                       datasource = DataSource,
                       pop = IcesAreaGroup,
                       pop.area.km2 = IcesAreaGroup.area.km2,
                       year = GrowingYear,
                       cohort = Cohort,
                       fyear = factor(GrowingYear),
                       fcohort = factor(Cohort),
                       pop.year = paste0(pop, ":", year),
                       pop.cohort = paste0(pop, ":", cohort)
                       # predictors
                       #f = FishingMortality,
                       #ssb.i = SpawningStockBiomass.1000t*1000/IcesAreaGroup.area.km2, #unit: ton/km2
                       )

# 6. keep only years with >= 10 increments/year for each pop  
data_otl <- data_otl %>% filter(pop.year %in% data_sum_10$pop.year)
```

## 2.2. Temperature data

```{r message=FALSE, warning=FALSE}
## load and process data
dir_temp <- "./data/temp"

# isimip
isimip <- read_rds(file.path(dir_temp, "isimip_sbt_datras_hist_ssp585.rds")) %>%
  mutate(date = Date,
         year = Year, 
         pop = if_else(IcesArea == "4abc", "4bc", IcesArea),
         temp = as.numeric(isimip_sbt),
         source = "isimip") %>%
  select(date:source) 

# oras5
oras <- read_rds(file.path(dir_temp, "oras5_datras.rds")) %>%
  mutate(date = Date,
         year = Year, 
         pop = if_else(IcesArea == "4abc", "4bc", IcesArea),
         temp = oras_sbt,
         source = "oras5") %>%
  select(date:source)

# nemo-medusa
nm <- read_rds(file.path(dir_temp, "nemomedusa_datras.rds")) %>%
  mutate(date = Date,
         year = as.numeric(Year), 
         pop = if_else(IcesArea == "4abc", "4bc", IcesArea),
         temp = nemomedusa_sbt,
         source = "nemo-medusa") %>%
  select(date:source)

## merge all temp data
# merge and summarize mean(temp) 
data_temp <- bind_rows(isimip, oras, nm) %>%
  filter(pop %in% c("4bc", "7a", "8ab"),
         year >= min(data_otl$year), year <= max(data_otl$year)) %>%
  group_by(source, pop, year) %>%
  summarize(temp = mean(temp))

# get centered temp (c.temp) for each pop
data_temp <- data_temp %>%
  group_by(source, pop) %>%
  mutate(c.temp = c.(temp))

# add source name
df_temp_name <- tibble(source = c("isimip", "oras5", "nemo-medusa"),
                       source_name = c("ISIMIP", "ORAS5", "NEMO-MEDUSA"))
data_temp <- data_temp %>% left_join(df_temp_name)

# add pop name
data_temp <- data_temp %>% left_join(df_pop)
```

## 2.3. Fishing data
```{r message=FALSE, warning=FALSE}
## Fishing data ----
# Fishing mortality, Spawning Stock Biomass, Recruitment from ICES Stock Assessment

dir_ices <- "./data/ices"

# sole distribution area - survey datras
datras <- read_sf(file.path(dir_ices, "hl_loc_4abc7a8ab.gpkg"))
datras <- as_tibble(datras) %>%
  mutate(pop = if_else(Area_27 == "4abc", "4bc", Area_27)) %>%
  select(pop, area_km2)

# sole stock assessment
data_sol <- read_rds(file.path(dir_ices, "stock-assessment_2023.rds")) %>%
  rename(ssb = SSB,
         f = `F`) %>%
  select(pop, year, ssb, f, recruitment) %>%
  left_join(datras) %>%
  mutate(ssb.i = ssb/area_km2,
         recruitment.i = recruitment/area_km2)

data_sol <- data_sol %>% left_join(df_pop)

# plaice stock assessment
data_ple <- read_rds(file.path(dir_ices, "ple_stock-assessment_2023.rds")) %>%
  rename(ssb_ple = SSB,
         recruitment_ple = recruitment) %>%
  select(pop, year, ssb_ple, recruitment_ple) %>%
  left_join(datras) %>%
  mutate(ssb_ple.i = ssb_ple/area_km2,
         recruitment_ple.i = recruitment_ple/area_km2)

```


## 2.4. Nutrient data

```{r message=FALSE, warning=FALSE}
## Nutrient data ----
dir_nu <- "./data/nutrient"
data_nu <- read_rds(file.path(dir_nu, "ospar_subset_1978-2017_ices_4abc.rds"))
# summarize all river by year 
data_nu <- data_nu %>% 
  group_by(IcesArea, year) %>% 
  summarize(TN = sum(TN)/1000,
            TP = sum(TP)/1000) %>%
  mutate(IcesArea = "4bc") %>%
  rename(pop = IcesArea)
# unit: 1000tN/year, 1000tP/year

data_nu <- data_nu %>% left_join(df_pop)
```


## 2.5. Model data

```{r}
#### BEST INTRINSIC STRUCTURE: 
m3 <- read_rds(file.path(dir_output, "intrinsic.model_best.rds"))

#### BEST EXTRINSIC STRUCTURE: 
## no scale
m4_isimip <- read_rds(file.path(dir_output, "extrinsic.model_best_isimip.rds"))
m4_oras5 <- read_rds(file.path(dir_output, "extrinsic.model_best_oras5.rds"))
m4_nm <- read_rds(file.path(dir_output, "extrinsic.model_best_nemo-medusa.rds"))
  
## scaled
m4_isimip_s <- read_rds(file.path(dir_output, "extrinsic.model_best_scaled_isimip.rds"))
m4_oras5_s <- read_rds(file.path(dir_output, "extrinsic.model_best_scaled_oras5.rds"))
m4_oras5_s <- read_rds(file.path(dir_output, "extrinsic.model_best_scaled_nemo-medusa.rds"))

#### BEST EXTRINSIC STRUCTURE WITH NUTRIENT DATA:
## no scale
m5_isimip <- read_rds(file.path(dir_output, "extrinsic.model.nu_best_isimip.rds"))
m5_oras5 <- read_rds(file.path(dir_output, "extrinsic.model.nu_best_oras5.rds"))
m5_nm <- read_rds(file.path(dir_output, "extrinsic.model.nu_best_nemo-medusa.rds"))
  
## scaled
m5_isimip_s <- read_rds(file.path(dir_output, "extrinsic.model.nu_best_scaled_isimip.rds"))
m5_oras5_s <- read_rds(file.path(dir_output, "extrinsic.model.nu_best_scaled_oras5.rds"))
m5_nm_s <- read_rds(file.path(dir_output, "extrinsic.model.nu_best_scaled_nemo-medusa.rds"))
```

# 3. DESCRIPTIVE RESULTS

## 3.1. Sampling size

### All data

```{r message=FALSE, warning=FALSE}
# Sampling size
## number of increment measurements
nrow(otl) 
## number of fish
n_distinct(otl$FishID) 
## range GrowingYear
range(otl$GrowingYear) 
## range Cohort
range(otl$Cohort) 
## range AgeAtCapture
range(otl$AgeAtCapture)

## number of fish per population
otl <- otl %>% left_join(df_pop)
otl %>% group_by(pop.name) %>% summarize(n_distinct(FishID))
```

```{r message=FALSE, warning=FALSE}
## Sampling distribution by Archive, Fish population, Range of collection year, Preparation method, and Reading institute (reader)

### add archive info
list_ifremer <- unique(filter(otl, str_detect(FishID, "RE|AL|CO") == TRUE)$FishID)
otl <- otl %>% mutate(archive = if_else(FishID %in% list_ifremer, "IFREMER", DataSource))


otl %>% 
  group_by(archive) %>% 
  summarize(method = unique(OtolithProcessingMethod))

otl %>% 
  group_by(archive) %>% 
  summarize(n_fish = n_distinct(FishID),
            sampling.year = paste0(min(SamplingYear), "-", max(SamplingYear)))  
otl %>% 
  group_by(archive, pop.name) %>% 
  summarize(pop = unique(pop.name),
            n_fish = n_distinct(FishID),
            sampling.year = paste0(min(SamplingYear), "-", max(SamplingYear)),
            Reader = unique(Reader)
            )  

```


```{r message=FALSE, warning=FALSE}
## AgeAtCapture distribution
otl_sum <- otl %>% group_by(pop.name, FishID, AgeAtCapture) %>% summarize(n = n()) %>%
  group_by(pop.name, AgeAtCapture) %>% summarize(n_fish = n())
ggplot(data = otl_sum, aes(x = AgeAtCapture, y = n_fish)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~ pop.name) +
  labs(x = "Age at capture (years)",
       y = "Number of fish")
```

```{r message=FALSE, warning=FALSE}
## sampling size vs SamplingYear/Cohort vs pop
otl_sum <- otl %>% group_by(pop.name, SamplingYear, Cohort, AgeAtCapture) %>% summarize(n_fish = n())

### SamplingYear
ggplot(data = otl_sum, aes(x = SamplingYear, y = AgeAtCapture, size = n_fish)) + 
  geom_point(alpha = 0.3) + 
  facet_wrap(~ pop.name) +
  labs(x = "Collection year",
       y = "Age at capture (years)",
       size = "Number of otolith") +
  scale_size(breaks = c(10, 50, 100, 250))

### Cohort
#ggplot(data = otl_sum, aes(x = Cohort, y = AgeAtCapture, size = n_fish)) + 
#  geom_point(alpha = 0.3) + 
#  facet_wrap(~ pop.name) 

```
### Analyzed data
```{r message=FALSE, warning=FALSE}
# Analyzed data
## number of increment measurements
nrow(data_otl) 
## number of fish
n_distinct(data_otl$FishID) 
## range GrowingYear
range(data_otl$GrowingYear) 
## range Cohort
range(data_otl$Cohort) 
## range AgeAtCapture
range(data_otl$AgeAtCapture)
```

## 3.2. Increment per year per population

```{r message=FALSE, warning=FALSE}
data_sum <- data_sum %>% left_join(df_pop)

ggplot() +
  geom_line(data = data_sum, 
            aes(x = GrowingYear, y = n_increment)) +
  geom_hline(yintercept = 10, linetype = "dashed") +
  #geom_hline(yintercept = 100, linetype = "dashed") +
  facet_wrap(~ pop.name) +
  labs(x = "Year",
       y = "Number of increment",
       color = NULL) +
  scale_y_continuous(breaks = c(0, 10, 100, 200, 300))
```

## 3.2. Otolith increments vs Extrinsic predictors

### Otolith width vs fish length
```{r message=FALSE, warning=FALSE}
# note - remove 2 fish with large size > 400mm but very small width < 1500, likely typo
otl_sub <- otl %>% 
  group_by(FishID) %>%
  summarize(width = max(OtolithWidth.um),
            fish.length = max(Length.mm)) %>%
  filter(width >= 1500)

nrow(otl_sub)

lm <- lm(fish.length ~ width, data = otl_sub)
summary(lm)
# Relationship between total fish length and otolith width (measured along the measurement axis on the otolith section) (Fish length ~ 51.42 + 0.076 * Otolith width, adjusted R2 = 0.53, p-value: < 0.001).

#cor.test(otl_sub$fish.length, otl_sub$width)

ggplot(data = otl_sub, aes(x = width, y = fish.length)) +
  geom_point(alpha = 0.3) +
  geom_smooth(color = "black", method = "lm") +
  labs(x = "Otolith width (μm)",
       y = "Total fish length (mm)")

```


### Otolith increments
```{r message=FALSE, warning=FALSE}
# otolith increment
data_otl <- data_otl %>% left_join(df_pop)

ggplot(data = data_otl, aes(x = Age, y = AnnulusDiameterIncrement.um, group = FishID)) + 
  geom_point(alpha = 0.3) +
  geom_line(alpha = 0.1) + 
  facet_grid(~ pop.name) +
  labs(x = "Age (years)",
       y = "Otolith increment (μm)")
```

### Extrinsic predictors
#### ssb, recruitment, f
```{r message=FALSE, warning=FALSE}
# sbt, ssb.i, f
## ssb.i
ggplot(data = data_sol %>% filter(year >= min(data_otl$year),
                                  year <= max(data_otl$year)), 
       aes(x = year, y = ssb.i, color = pop.name)) + 
  geom_line() +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Year",
       y = "Spawning Stock Biomass (tonne/km²)",
       color = NULL)

## recruitment.i
ggplot(data = data_sol %>% filter(year >= min(data_otl$year),
                                  year <= max(data_otl$year)),
       aes(x = year, y = recruitment.i, color = pop.name)) + 
  geom_line() +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Year",
       y = "Recruitment (thousands/km²)",
       color = NULL)

## f
ggplot(data = data_sol %>% filter(year >= min(data_otl$year),
                                  year <= max(data_otl$year)),
       aes(x = year, y = f, color = pop.name)) + 
  geom_line() +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Year",
       y = "Fishing Mortality",
       color = NULL)
```

#### Nutrient - TN, TP
```{r message=FALSE, warning=FALSE}
## TN
ggplot(data = data_nu %>% filter(year >= min(data_otl$year),
                                  year <= max(data_otl$year)),
       aes(x = year, y = TN, color = pop.name)) + 
  geom_line() +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Year",
       y = "Total Nitrogen (kilotonne/year)",
       color = NULL) +
  scale_color_manual(values = "#7570B3")

## TP
ggplot(data = data_nu %>% filter(year >= min(data_otl$year),
                                  year <= max(data_otl$year)),
       aes(x = year, y = TP, color = pop.name)) + 
  geom_line() +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Year",
       y = "Total Phosphorus (kilotonne/year)",
       color = NULL) +
  scale_color_manual(values = "#7570B3")

## TN and TP
ggplot(data = data_nu %>% filter(year >= min(data_otl$year),
                                  year <= max(data_otl$year))) +
  geom_line(aes(x = year, y = scale(TN), color = pop.name, linetype = "Total Nitrogen")) +
  geom_line(aes(x = year, y = scale(TP), color = pop.name, linetype = "Total Phosphorus")) +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Year",
       y = "Scaled value",
       color = NULL,
       linetype = NULL) +
  scale_color_manual(values = "#7570B3")
```


#### Temperature

potential sea temperature at sea floor: isimip, oras5, nemo-medusa

##### monthly variation

monthly variation of temperature is quite similar across datasets, except for hadisst in the Bay of Biscay. Sea surface temperature hadisst is always higher than potential sea temperature at sea floor datasets.

```{r message=FALSE, warning=FALSE}
#### data
data_temp_month <- bind_rows(isimip, oras, nm) %>%
  filter(pop %in% c("4bc", "7a", "8ab"),
         year >= min(data_otl$year), year <= max(data_otl$year)) %>%
  mutate(month = month(date)) %>%
  group_by(source, pop, month) %>%
  summarize(temp = mean(temp)) %>%
  left_join(df_pop) %>%
  left_join(df_temp_name)

#### plot
ggplot(data = data_temp_month, aes(x = month, y = temp, color = source_name)) +
  geom_line() +
  facet_grid(. ~ pop.name) +
  scale_x_continuous(breaks = c(3, 6, 9, 12)) +
  labs(x = "Month",
       y = "Temperature (°C)",
       color = "Dataset") +
  scale_colour_manual(values = c("#7CAE00","#F8766D", "#00BFC4")) +
  scale_fill_manual(values = c("#7CAE00","#F8766D", "#00BFC4")) 
```

##### yearly variation (yearly temperature - mean temperature)

```{r}
#### isimip, nemo-medusa, oras5 (reprenting: hadisst)
ggplot() +
  geom_line(data = data_temp, 
       aes(x = year, y = c.temp, color = source_name)) +
  # geom_vline(data = data_temp, aes(xintercept = min(nm$year)), linetype = "dashed", alpha = 0.5) +
  facet_grid(pop.name ~ .) +
  labs(x = "Year",
       y = "Temperature variation (°C)",
       color = "Dataset") +
  scale_colour_manual(values = c("#7CAE00","#F8766D", "#00BFC4")) +
  scale_fill_manual(values = c("#7CAE00","#F8766D", "#00BFC4")) 

```