---
title: "si_otolith reading precision"
output: html_document
date: "2023-07-13"
bibliography: _si_references.bib
csl: _ices-journal-of-marine-science.csl
---

**BEFORE:**

-   fwo_phd_git/Data/Env/isimip_thetao_process.R

-   fwo_phd_git/Data/otolith/2_prepare data for analysis.R

**AFTER:**

# SETUP

```{r message=FALSE, warning=FALSE}
# PACKAGES
library(tidyverse)    # process data
library(FSA)          # precision calculation

# WORKING DIRECTORY
dir_otl <- "D:/OneDrive - UGent/data/WP1/otolith"

# NOTE
# reference: https://derekogle.com/fishR/examples/oldFishRVignettes/AgeComparisons.pdf
```

# AGING PRECISION - ALL SAMPLES

This sections calculate the aging precision of the sampled otoliths between less experienced readers Tuan Anh Bui (tbui/TAB), Kelly Diaz (kdiaz/KD) and a experienced reader from ILVO Ilse Maertens (imaertens).

```{r}
# setup
otl_reage <- read_rds(file.path(dir_otl, "otl_ilvo_reage.rds"))
otl_ilvo <- read_rds(file.path(dir_otl, "otl_ilvo.rds")) #data used for analysis
otl_ilvo <- otl_ilvo %>% filter(FishID != "SOL_52_G1_Z.121_26-03-1990_1602")

otl <- otl_reage %>% 
  filter(FishID %in% otl_ilvo$FishID) %>%
  filter(FishID != "SOL_52_G1_Z.121_26-03-1990_1602") %>%
  select(FishID, Reader, AgeAtCapture) %>% 
  unique() %>%
  pivot_wider(names_from = Reader, values_from = AgeAtCapture)

# percentage re-aged
n_read <- sum(!is.na(otl$imaertens))/n_distinct(otl$FishID)*100 #99.5% otoliths re-age
```

```{r}
# tbui vs imaertens
otl_tbui <- otl %>% 
  select(-kdiaz) %>%
  filter(!is.na(tbui))

ap_tbui <- agePrecision(~ tbui + imaertens, data = otl_tbui)
#summary(ap_tbui, what="absolute difference")
summary(ap_tbui, what="precision")
```

```{r}
# kdiaz vs imaertens
otl_kdiaz <- otl %>% 
  select(-tbui) %>%
  filter(!is.na(kdiaz))

ap_kdiaz <- agePrecision(~ kdiaz + imaertens, data = otl_kdiaz)
#summary(ap_kdiaz,what="absolute difference")
summary(ap_kdiaz, what="precision")
```

`r round(n_read,1)`% of all sampled otoliths was re-aged. 

Coefficient of variation between tbui and imaertens is `r round(summary(ap_tbui, what="precision")$ACV,1)`%.

Coefficient of variation between kdiaz and imaertens is `r round(summary(ap_kdiaz, what="precision")$ACV,1)`%.

The aging is considered precised as the coefficient of variations are under 5% [@Campana2001]

# AGING AND MEASURING PRECISION - BAY OF BISCAY

```{r}
# setup
otl_8ab <- read_rds(file.path(dir_otl, "otl_ilvo_consistency.rds"))
otl_8ab <- otl_8ab %>% filter(FishID != "SOL_52_G1_Z.121_26-03-1990_1602")

# percentage re-aged
n_read_8ab <- n_distinct(otl_8ab$FishID)/n_distinct(filter(otl_ilvo, Reader == "kdiaz")$FishID)*100 #16.4%% otoliths re-age
```

```{r}
# aging precision
otl_age <- otl_8ab %>% 
  select(FishID, Reader, AgeAtCapture) %>%
  unique() %>%
  pivot_wider(names_from = Reader, values_from = AgeAtCapture) 

ap_age <- agePrecision(~ tbui + kdiaz, data = otl_age)
summary(ap_age, what="precision")
```

```{r}
# measuring precision
otl_width <- otl_8ab %>% 
  select(FishID, Reader, Age, AnnulusDiameter.um) %>%
  pivot_wider(names_from = Reader, values_from = AnnulusDiameter.um) 

ap_width <- agePrecision(~ tbui + kdiaz, data = otl_width)
summary(ap_width, what="precision")
```


`r round(n_read_8ab,1)`% of Bay of Biscay otoliths was re-aged and re-measured. 

Coefficient of variation of aging precision is `r round(summary(ap_age, what="precision")$ACV,1)`%.

Coefficient of variation of measuring precision is `r round(summary(ap_width, what="precision")$ACV,1)`%.

The aging and measuring are considered precised as the coefficient of variations are under 5% [@Campana2001]

# REFERENCES
